On Supporting Molecular Visualization on Grid Resources using Visual Programming Environments

### 2008-04-24 opal/documents/grid2008/paper.tex
added a couple of diagrams and some text - more coming

\section {Background}

@@

\section {Application based services}

\subsection {Opal Overview}

- Opal wrapper and GUI \cite{opal}: Sriram 
- Opal and Opal-CSF4: Sriram
- Data services: Sriram

\subsection {Towards Opal2}

>>>>

\section {Visual Programming Environments}

@@

\section {Services Oriented Science}

Programming and using distributed Grid resources is quite a challenge task.
There is hope that Service-Oriented Architectures (SOA) can address this
problem by providing platform-independent, language-neutral service
interfaces that hide the complexity of the implementations while providing
a well-defined and high-performance Quality of Service (QoS). In our recent
paper [ref], we claimed that in order to lower the entrance barrier for
scientific application developers and end-users to Grid infrastructures, we
must do the following:

% TODO: possibly expand every bullet
\begin{itemize}
\item{Focus on application-level services}
\item{Provide users access to virtualized resources}
\item{Provide multiple user interfaces to scientific users}
\item{Lower deployment costs}
\item{Follow the Web 2.0 paradigm by enabling user participation and contribution}
\end{itemize}

Towards this end, we have developed the Opal toolkit
\cite{opal} for wrapping scientific applications as Web services. In order
to wrap an application as a Web service, application developers are
expected to write a simple XML-based application configuration. The
application configuration contains information about the scientific
application, such as application name, binary location, and application
metadata (e.g. usage information). Using a simple Apache Ant task, the Opal
toolkit can deploy the application as a Web service into a container based
on Apache Tomcat and Axis. Once the application is deployed as a Web
service, clients can access this service programmatically using its WSDL
description. The WSDL API provides operations for job launch (which accepts
command-line arguments and input files as its parameters), querying status,
and retrieving outputs. Furthermore, it also provides an API for retrieving
application metadata. WSDL savvy users could write their own clients to
access Opal based applications. However, not every scientific user is
capable or interested in writing Web service clients. Hence, several
interfaces have been provided for the end-user. Apart from command-line
clients written in languages like Java, Python [4] and Perl [15] available
from NBCR, Web interfaces for Opal based services are available through the
GridSphere portal environment [18], such as the MEME portlet in My
WorkSphere [24]. In addition, Opal based interfaces via Rich Internet
Applications (RIA) such as Gemstone [20], and Problem Solving Environments
(PSE) such as the Python Molecular Viewer (PMV) [22], AutoDockTools [21]
and Continuity [27] have also been provided.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{architecture}
\end{center}
\caption{Services Oriented Science Using the Opal Toolkit}
\label{fig:arch}
\end{figure}

Although a detailed description of the Opal toolkit is beyond the scope of
this paper (and can be found in \cite{opal}), the overall architecture of
is shown in Figure \ref{fig:arch}. The salient features of Opal are as
follows:

\begin{itemize}
\item{Scheduling and Cluster Management}
\item{Data management and Persistence}
\item{Security}
\item{Standardized WSDL API}
\item{Argument Description and Dynamic Interface Generation}
\end{itemize}

- Opal wrapper and GUI: Sriram

- Opal and Opal-CSF4: Sriram

- Data services: Sriram

- Status, more information

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{opal-gui}
\end{center}
\caption{Service Invocation via the Opal GUI}
\label{fig:opal-gui}
\end{figure}

- Towards Opal2

\begin{figure}
\begin{center}
\includegraphics[width=8.5cm]{opal2-providers}
\end{center}
\caption{The Opal2 Provider Model}
\label{fig:opal2-providers}
\end{figure}
